Drug sensitivity prediction with normal inverse Gaussian shrinkage informed by external data

### 2020-06-18 docs/manuscript.Rnw
removed Makevars files from Code_and_Data

  To summarise, estimation of only $\bm{\alpha}_{\text{feat}}$ or
  $\bm{\alpha}_{\text{drug}}$ (and consequently $\phi_{jd}$ and $\chi_{d}$)
  by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively is relatively unbiased,
  as evident from simulation Scenarios 1 and 2 (Figures 
  \ref{fig:simulations_gdsc_est1} and \ref{fig:simulations_gdsc_est2}). 
  In contrast, simultaneous 
  estimation in Scenario 3 results in overestimated $\phi_{jd}$ and 
  underestimated $\chi_{d}$ by NIG$_{\text{f}+\text{d}}$
  (Figure \ref{fig:simulations_gdsc_est3}). We conjecture that this interplay
  of drug and feature variance components is due to near-unidentifiability.e
  r, the resultconsequences are limited, since theiances on the $\beta_{jd}$ level are unbiasleft ed 
  (Figure \ref{fig:simulations_gdsc_var}). If separately estimated, scale 
  parameters $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are 
  overestimated by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively.
  If estimated simultaneously NIG$_{\text{f}+\text{d}}$ overestimates 
  $\lambda_{\text{feat}}$ and underestimated $\lambda_{\text{drug}}$. On the
  $\beta_{jd}$ level, $\bm{\lambda}$ influences kurtoses 
  $\mathcal{K}(\beta_{jd})$. Figure \ref{fig:simulations_gdsc_kurt} shows the
  mean estimated kurtoses versus their true values. Kurtoses seem to be
  underestimated by NIG$_{\text{f}+\text{d}}$. In contrast, 
  NIG$_{\text{f}+\text{d}}^{-1}$ overestimates the true mean of the
  $\mathcal{K}(\beta_{jd})$, \Sexpr{true.kurtb3}, at \Sexpr{kurtb3}.
  Lastly, the results from Scenario 3 in SM Section \ref{sm-sec:simulations} 
  show that the VB approximation is quite good as compared to standard MCMC.
<<simulations_gdsc_kurt, fig.cap="Simulation results for Scenario 3: mean estimated prior Kurtoses $\\mathcal{K}(\\beta_{jd})$ versus true values, with line of identity (dotted).", out.width="50%", fig.asp=1>>= 
@

@@

tab1 <- t(fit1)[c(1, 2), c(1:3)]
rownames(tab1) <- c("NIG$_{\\text{f}+\\text{d}}^-$", 
                    "NIG$_{\\text{f}+\\text{d}}$")
colnames(tab1) <- c("feature intercept", "$\\log p$-value", "drug intercept")

@@

tab3 <- t(fit3)[c(1, 2), c(1:3)]
rownames(tab3) <- c("NIG$_{\\text{f}+\\text{d}}^-$", 
                    "NIG$_{\\text{f}+\\text{d}}$")
colnames(tab3) <- c("feature intercept", "$\\log p$-value", "drug intercept")

@@

tab5.2 <- tab5[, c(4, 13:14, 10:12)]
tabm5.2 <- tabm5[, c(4, 13:14, 10:12)]
tab5.2 <- t(sapply(1:nrow(tab5.2), function(r) {
  id <- which(tabm5.2[r, ]==min(tabm5.2[r, ], na.rm=TRUE))
  out <- tab5.2[r, ]
  out[id] <- paste0("\\textbf{", out[id], "}")
  out}))
rownames(tab5.2) <- paste0("analysis ", 1:nrow(tab5.2))
colnames(tab5.2) <- c("lasso($\\sim$50)$^*$", "lasso(25)", "lasso(100)", 
                      "NIG$_{\\text{f}+\\text{d}}+$DSS($\\sim$50)$^*$",
                      "NIG$_{\\text{f}+\\text{d}}+$DSS(25)",
                      "NIG$_{\\text{f}+\\text{d}}+$DSS(100)")
tab5.2 <- t(tab5.2)
tab5.2 <- tab5.2[c(2, 1, 3, 5, 4, 6), ]

@@

# pmse ranges
ranges1 <- apply(rbind(mpmse1, mpmse2, mpmse3, mpmse4)[, 1:4], 1, range) 
ranges1 <- round(ranges1[2, ] - ranges1[1, ], 4)
psels1 <- round(mpsel[, c(4, 6, 10)], 0)
@
  The non-zero NIG$_{\text{f}+\text{d}}$ $\bm{\alpha}$ estimates in Tables 
  \ref{tab:analysis_gdsc_fit1}-\ref{tab:analysis_gdsc_fit4.2} show 
  that there is an effect of the
  external covariates. 
  \begin{itemize}
	  \item{Analysis 1 results in a positive additive effect of the mutation 
	    $log p$-values on the prior $\bm{\beta}_d$ precisions
	    (Table \ref{tab:analysis_gdsc_fit1}). More precisely, with doubling
	    $p$-value, the prior $\bm{\beta}_d$ precision increases by 
	    $\log(2)*\Sexpr{round(tab1[2, 2], 3)}$=\Sexpr{
	    round(log(2)*tab1[2, 2], 3)}.
	    This translates to more $\bm{\beta}_d$ shrinkage with increasing 
	    $p$-value, as expected.}
	  \item{Analysis 2 shows that gene mutations are more predictive 
	    than gene expressions, as observed from the negative effect of mutation
	    dummy on prior precisions (Table \ref{tab:analysis_gdsc_fit2}): 
	    mutations are shrunken less than
	    expressions.}
	  \item{Analysis 3 indicates that CCLE $\log p$-values are positively related
	    to prior precision (Table \ref{tab:analysis_gdsc_fit3}). In particular
	    if the $p$-value doubles, the prior $\bm{\beta}_d$ precision increases by 
	    $\log(2)*\Sexpr{round(tab3[2, 2], 3)}$=\Sexpr{
	    round(log(2)*tab1[2, 2], 3)}, i.e., higher CCLE $p$-values results in 
	    more shrinkage of $\bm{\beta}_d$, as expected.}
	  \item{Analysis 4 gives a negative effect for the pathway dummy (Table 
	    \ref{tab:analysis_gdsc_fit4.1}), indicating less shrinkage for genes
	    that are in the drugs' target pathway, as expected. According to 
	    expectation, experimental and 
	    developmental drugs prior precisions are shrunken more than
	    the reference category, approved drugs (Table 
	    \ref{tab:analysis_gdsc_fit4.2}). Somewhat surprisingly, targeted and 
	    unkown drugs are shrunken more than the reference, cytotoxic drugs (Table 
	    \ref{tab:analysis_gdsc_fit4.2}), although the differences on the
	    $\bm{\beta}_d$ variance scale are small:
	    $\hat{\E}(\tau_{d}^{2}) = \{ \Sexpr{vars1[1]}, \Sexpr{vars1[2]}, 
	    \Sexpr{vars1[3]} \}$ for cytotoxic, targeted, and unkown drugs, 
	    respectively (ignoring other variance components).}
	\end{itemize}
<<analysis_gdsc_fit1>>=
kableExtra::kable_styling(knitr::kable(
  tab1, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 1. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
<<analysis_gdsc_fit2>>=
kableExtra::kable_styling(knitr::kable(
  tab2, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 2. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
<<analysis_gdsc_fit3>>=
kableExtra::kable_styling(knitr::kable(
  tab3, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 3. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
  \begin{table}[h!]\label{tab:analysis_gdsc_fit4}
    \begin{minipage}{1\textwidth}
<<analysis_gdsc_fit4.1>>=
kableExtra::kable_styling(knitr::kable(
  tab4.1, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 4. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
    \end{minipage}
    \begin{minipage}{1\textwidth}
<<analysis_gdsc_fit4.2>>=
kableExtra::kable_styling(knitr::kable(
  tab4.2, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 4. Empty cells correspond to fixed zero parameters \\textit{(continued)}.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
    \end{minipage}
  \end{table}

  The mean prediction mean squared errors (PMSE), calculated on the 
  test data, are diplayed in Table \ref{tab:analysis_gdsc_pmse1}.
  Note that due to standardisation, an empty reference model has a PMSE of one.
  In terms of PMSE, ridge outperforms the other models in all Analyses 1, 3, and
  4, while NIG$_{\text{f}+\text{d}}$ $\bm{\alpha}$ performs best in Analysis 2. 
  In general however, all models perform very similar. The ranges of mean PMSE 
  over all methods are
  \Sexpr{ranges1[1:3]}, and \Sexpr{ranges1[4]}, for Analyses 1-4, respectively,
  indicating that the differences in predictive performance are very small.
  Furthermore, the difference between NIG$_{\text{f}+\text{d}}$, that includes
  external covariates, and NIG$_{\text{f}+\text{d}}^-$,
  that excludes the external covariates is small; an indication that the
  external covariates are not very informative here.
<<analysis_gdsc_pmse1>>=  
kableExtra::kable_styling(knitr::kable(
  tab5.1, align="r",
  caption="Mean (standard deviation) of cross-validated PMSE for GDSC data. Best performing model (per analysis) in bold.",
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

@@

<<analysis_gdsc_pmse2>>=  
kableExtra::add_footnote(kableExtra::kable_styling(knitr::kable(
  tab5.2, align="r",
  caption="Mean (standard deviation) of cross-validated PMSE for selection methods (number of selected features between parentheses) on GDSC data. Best performing model (per analysis) in bold.",
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position")), 
  label="cross validated lasso selects, on average, 42-56 features in all analyses (indicated with $\\sim$50).",
  notation="symbol", threeparttable=TRUE, escape=FALSE)
@

@@

	An alternative strategy to include external covariates is the varying 
	coefficient (VC) model \cite[]{hastie_varying-coefficient_1993}. The VC model
	treats the mean of the regression coefficients as a deterministic function of 
	external covariates, as opposed to our probabilistic model for the variance
	of the regression coefficients. \cite{ni_bayesian_2019} introduces a Bayesian 
	VC model where the relation between the regression 
	coefficients and external covariates is no longer deterministic, but still
	based on the mean of the coefficients. Besides our computationally more
	feasible VB-EM estimation, we advocate for a more indirect
	model for the relation between external covariates and regression 
	coefficients, i.e., through their random variance components. This indirect 
	model assumes less structure about the relation between external covariates
	and regression coefficients than the direct VC mean model approach.
  In particular, a VC mean model describes both magnitude and direction of the 
  external covariates effects, while our variance model only describes the
  magnitude and is invariant to the direction of the effects.
  Nonetheless, a combination of the two approaches (both mean and variance 
  modelled as functions of the external covariates) may be a fruitful future 
  research direction.
  Other methods that consider an external covariate model for the variance
  components of the regression coefficients are bSEM
  \cite[]{leday_gene_2017,kpogbezan_empirical_2017} and \texttt{xtune}
  \cite[]{zeng_incorporating_2020}. bSEM is designed to include
  only dichotomous external covariates, so is not applicable in most of the
  applications and simulations that we have considered here. \texttt{xtune}
  allows for the inclusion of multiple external feature covariates, but not on
  the drug level, so also has limited applicability in our simulations and
  applications.

@@

	Possible directions of future research are applications of the NIG model to 
	different data types. Suitable applications are eQTL studies, 
	in which gene expressions are regressed on SNPs. Several interesting external
	covariates are available, both on the genes as well as the SNPs. An example of
	such an external
	covariate for the genes is genelength, where we suspect that longer genes are
	harder to predict. The distance of the SNP to the gene is an example of an
	external SNP covariate, where the expectation is that SNPs further from the 
	gene are less predictive of that gene's expression.

>>>>

  To summarise, estimation of only $\bm{\alpha}_{\text{feat}}$ or
  $\bm{\alpha}_{\text{drug}}$ (and consequently $\phi_{jd}$ and $\chi_{d}$)
  by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively is relatively unbiased,
  as evident from simulation Scenarios 1 and 2 (Figures 
  \ref{fig:simulations_gdsc_est1} and \ref{fig:simulations_gdsc_est2}). 
  In contrast, simultaneous 
  estimation in Scenario 3 results in overestimated $\phi_{jd}$ and 
  underestimated $\chi_{d}$ by NIG$_{\text{f}+\text{d}}$
  (Figure \ref{fig:simulations_gdsc_est3}). We conjecture that this interplay
  of drug and feature variance components is due to near-unidentifiability. 
  In any case, the consequences are limited, since the variances on the 
  $\beta_{jd}$ level are left unbiased
  (Figure \ref{fig:simulations_gdsc_var}). If separately estimated, scale 
  parameters $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are 
  overestimated by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively.
  If estimated simultaneously NIG$_{\text{f}+\text{d}}$ overestimates 
  $\lambda_{\text{feat}}$ and underestimated $\lambda_{\text{drug}}$. On the
  $\beta_{jd}$ level, $\bm{\lambda}$ influences kurtoses 
  $\mathcal{K}(\beta_{jd})$. Figure \ref{fig:simulations_gdsc_kurt} shows the
  mean estimated kurtoses versus their true values. Kurtoses seem to be
  underestimated by NIG$_{\text{f}+\text{d}}$. In contrast, 
  NIG$_{\text{f}+\text{d}}^{-1}$ overestimates the true mean of the
  $\mathcal{K}(\beta_{jd})$, \Sexpr{true.kurtb3}, at \Sexpr{kurtb3}.
  Lastly, the results from Scenario 3 in SM Section \ref{sm-sec:simulations} 
  show that the VB approximation is quite good as compared to standard MCMC.
<<simulations_gdsc_kurt, fig.cap="Simulation results for Scenario 3: mean estimated prior Kurtoses $\\mathcal{K}(\\beta_{jd})$ versus true values, with line of identity (dotted).", out.width="50%", fig.asp=1>>= 
@

@@

tab1 <- t(fit1)[c(1, 2), c(1:3)]
rownames(tab1) <- c("NIG$_{\\text{f}+\\text{d}}^-$", 
                    "NIG$_{\\text{f}+\\text{d}}$")
colnames(tab1) <- c("feature intercept", "$p$-value", "drug intercept")

@@

tab3 <- t(fit3)[c(1, 2), c(1:3)]
rownames(tab3) <- c("NIG$_{\\text{f}+\\text{d}}^-$", 
                    "NIG$_{\\text{f}+\\text{d}}$")
colnames(tab3) <- c("feature intercept", "$p$-value", "drug intercept")

@@

tab5.2 <- tab5[, c(4, 13:14, 10:12)]
tabm5.2 <- tabm5[, c(4, 13:14, 10:12)]
tab5.2 <- t(sapply(1:nrow(tab5.2), function(r) {
  id <- which(tabm5.2[r, ]==min(tabm5.2[r, ], na.rm=TRUE))
  out <- tab5.2[r, ]
  out[id] <- paste0("\\textbf{", out[id], "}")
  out}))
rownames(tab5.2) <- paste0("analysis ", 1:nrow(tab5.2))
colnames(tab5.2) <- c("lasso$_{\\lambda^{\\text{CV}}}$($\\sim$50)$^*$", 
                      "lasso(25)", "lasso(100)", 
                      "NIG$_{\\text{f}+\\text{d}}+$DSS($\\sim$50)$^*$",
                      "NIG$_{\\text{f}+\\text{d}}+$DSS(25)",
                      "NIG$_{\\text{f}+\\text{d}}+$DSS(100)")
tab5.2 <- t(tab5.2)
tab5.2 <- tab5.2[c(2, 1, 3, 5, 4, 6), ]

@@

# pmse ranges
ranges1 <- apply(rbind(mpmse1, mpmse2, mpmse3, mpmse4)[, 1:4], 1, range) 
ranges1 <- round(ranges1[2, ] - ranges1[1, ], 4)
psels1 <- round(mpsel[, c(4, 6, 10)], 0)
@
  The non-zero NIG$_{\text{f}+\text{d}}$ $\bm{\alpha}$ estimates in Tables 
  \ref{tab:analysis_gdsc_fit1}-\ref{tab:analysis_gdsc_fit4.2} show 
  that there is an effect of the
  external covariates. 
  \begin{itemize}
	  \item{Analysis 1 results in a positive additive effect of the mutation 
	    $p$-values on the prior $\bm{\beta}_d$ precisions
	    (Table \ref{tab:analysis_gdsc_fit1}). 
	    This translates to more $\bm{\beta}_d$ shrinkage towards zero with 
	    increasing $p$-value, as expected.}
	  \item{Analysis 2 shows that gene mutations are more predictive 
	    than gene expressions, as observed from the negative effect of mutation
	    dummy on prior precisions (Table \ref{tab:analysis_gdsc_fit2}): 
	    mutations are shrunken less than
	    expressions.}
	  \item{Analysis 3 indicates that CCLE $p$-values are positively related
	    to prior precision (Table \ref{tab:analysis_gdsc_fit3}), i.e., higher 
	    CCLE $p$-values results in 
	    more shrinkage of $\bm{\beta}_d$, as expected.}
	  \item{Analysis 4 gives a negative effect for the pathway dummy (Table 
	    \ref{tab:analysis_gdsc_fit4.1}), indicating less shrinkage for genes
	    that are in the drugs' target pathway, as expected. According to 
	    expectation, experimental and 
	    developmental drugs prior precisions are shrunken more than
	    the reference category, approved drugs (Table 
	    \ref{tab:analysis_gdsc_fit4.2}). Somewhat surprisingly, targeted and 
	    unkown drugs are shrunken more than the reference, cytotoxic drugs (Table 
	    \ref{tab:analysis_gdsc_fit4.2}), although the differences on the
	    $\bm{\beta}_d$ variance scale are small:
	    $\hat{\E}(\tau_{d}^{2}) = \{ \Sexpr{vars1[1]}, \Sexpr{vars1[2]}, 
	    \Sexpr{vars1[3]} \}$ for cytotoxic, targeted, and unkown drugs, 
	    respectively (ignoring other variance components).}
	\end{itemize}
<<analysis_gdsc_fit1>>=
kableExtra::kable_styling(knitr::kable(
  tab1, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 1. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
<<analysis_gdsc_fit2>>=
kableExtra::kable_styling(knitr::kable(
  tab2, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 2. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
<<analysis_gdsc_fit3>>=
kableExtra::kable_styling(knitr::kable(
  tab3, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 3. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
  \begin{table}[h!]\label{tab:analysis_gdsc_fit4}
    \begin{minipage}{1\textwidth}
<<analysis_gdsc_fit4.1>>=
kableExtra::kable_styling(knitr::kable(
  tab4.1, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 4. Empty cells correspond to fixed zero parameters.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
    \end{minipage}
    \begin{minipage}{1\textwidth}
<<analysis_gdsc_fit4.2>>=
kableExtra::kable_styling(knitr::kable(
  tab4.2, align="r", digits=3,
  caption="$\\bm{\\alpha}$ estimates from analysis 4. Empty cells correspond to fixed zero parameters \\textit{(continued)}.",
  format="latex", booktabs=TRUE, escape=FALSE),
  latex_options=c("HOLD_position"))
@
    \end{minipage}
  \end{table}

  The mean prediction mean squared errors (PMSE), calculated on the 
  test data, are diplayed in Table \ref{tab:analysis_gdsc_pmse1}.
  Note that due to standardisation, an empty reference model has a PMSE of one.
  In terms of PMSE, ridge outperforms the other models in all Analyses 1, 3, and
  4, while NIG$_{\text{f}+\text{d}}$ $\bm{\alpha}$ performs best in Analysis 2. 
  In general however, all models perform very similarly. The ranges of mean PMSE 
  over all methods are
  \Sexpr{ranges1[1:3]}, and \Sexpr{ranges1[4]}, for Analyses 1-4, respectively,
  indicating that the differences in predictive performance are very small.
  Furthermore, the difference between NIG$_{\text{f}+\text{d}}$, that includes
  external covariates, and NIG$_{\text{f}+\text{d}}^-$,
  that excludes the external covariates is small; an indication that the
  external covariates are not very informative here.
<<analysis_gdsc_pmse1>>=  
kableExtra::kable_styling(knitr::kable(
  tab5.1, align="r",
  caption="Mean (standard deviation) of cross-validated PMSE for GDSC data. Best performing model (per analysis) in bold.",
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

@@

<<analysis_gdsc_pmse2>>=  
kableExtra::add_footnote(kableExtra::kable_styling(knitr::kable(
  tab5.2, align="r",
  caption="Mean (standard deviation) of cross-validated PMSE for selection methods (number of selected features between parentheses) on GDSC data. Best performing model (per analysis) in bold.",
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position")), 
  label="lasso with cross validated $\\lambda$ selects, on average, 42-56 features in all analyses (indicated with $\\sim$50).",
  notation="symbol", threeparttable=TRUE, escape=FALSE)
@

@@

	An alternative strategy to include external covariates is the varying 
	coefficient (VC) model \cite[]{hastie_varying-coefficient_1993}. The VC model
	treats the mean of the regression coefficients as a deterministic function of 
	external covariates, as opposed to our probabilistic model for the variance
	of the regression coefficients. \cite{ni_bayesian_2019} introduces a Bayesian 
	VC model where the relation between the regression 
	coefficients and external covariates is no longer deterministic, but still
	based on the mean of the coefficients. Besides our computationally more
	feasible VB-EM estimation, we advocate for a more indirect
	model for the relation between external covariates and regression 
	coefficients, i.e., through their random variance components. This indirect 
	model assumes less structure about the relation between external covariates
	and regression coefficients than the direct VC mean model approach.
  In particular, a VC mean model describes both magnitude and direction of the 
  external covariates effects, while our variance model only describes the
  magnitude and is invariant to the direction of the effects.
  Nonetheless, a combination of the two approaches (both mean and variance 
  modelled as functions of the external covariates) may be a fruitful future 
  research direction.
  Other methods that consider an external covariate model for the variance
  components of the regression coefficients are bSEM
  \cite[]{leday_gene_2017,kpogbezan_empirical_2017} and \texttt{xtune}
  \cite[]{zeng_incorporating_2020}. Here, bSEM is designed to include
  only dichotomous external covariates, so is not applicable in most of the
  applications and simulations that we have considered here. Inclusion of
  multiple external features is allowed by \texttt{xtune}, but not on
  the drug level, so \texttt{xtune} has limited applicability in our simulations 
  and applications.

@@

	Possible directions of future research are applications of the NIG model to 
	different data types. Suitable applications are eQTL studies, 
	in which gene expressions are regressed on SNPs. Several interesting external
	covariates are available, both on the genes as well as the SNPs. An example of
	such an external
	covariate for the genes is gene length, where we suspect that longer genes are
	harder to predict. The distance of the SNP to the gene is an example of an
	external SNP covariate, where the expectation is that SNPs further from the 
	gene are less predictive of that gene's expression.

### 2020-06-18 docs/supplement.Rnw
removed Makevars files from Code_and_Data

n_eff.mean <- round(mean(summary.mcmc2$summary[, "n_eff"]), 0)
n_eff.min <- round(min(summary.mcmc2$summary[, "n_eff"]), 0)
time.mean <- round((time.mcmc2*1000/n_eff.mean)/3600, 0)
time.min <- round((time.mcmc2*1000/n_eff.min)/(7*24*3600), 0)
@	
  The lasso and ridge models in this section were fit separately per drug, 
  using the \texttt{R} package \texttt{glmnet} 
  \cite[]{friedman_regularization_2010} with cross
	validated penalty parameters.

@@

  For completeness, Figure \ref{fig:simulations_gdsc_res5} displays samples
  from the full Bayesian NIG posterior, as introduced in Section 
  \ref{sec:gibbssampler}. Unfortunately, we were not able to generate enough
  samples for a reliable posterior using the Gibbs procedure as described in 
  the Section. We generated the samples from the full Bayesian posterior using
  stan through the R package \texttt{rstan} 
  \cite[]{guo_rstan:_2018} (see \url{https://github.com/magnusmunch/NIG} for 
  the implemented stan code). The Figure shows that the posteriors for
  $\bm{\phi}$ and $\bm{\chi}$ have unexpected shapes and seem to have explored
  only a small part of the posterior space. This indication that
  posterior inference is not reliable is confirmed by the average and minimum
  effective sample sizes of \Sexpr{n_eff.mean} and \Sexpr{n_eff.min}over all 
  parameters, respectively. For reliable inference,
  \cite{kruschke_doing_2015} recommends an effective sample size of at least 
  10 000. Inspection of the computation times and a quick calculation learns 
  that we have to sample from the posterior
  for at least \Sexpr{time.mean} hours to obtain an average effective sample 
  size of 10 000. Moreover, to achieve a minimum effective sample size of 10 000 
  for all parameters, the sampler has to be run for about \Sexpr{time.min}
  weeks.
<<simulations_gdsc_res5, fig.cap="Simulation results for Scenario 3: Samples from full Bayesian NIG posteriors (see Section \\ref{sec:gibbssampler}) with true value superimposed (red line) for (top row) group 1, (second row) group 2, (third row) group 3, and (bottom row) group 4 for both features and drugs, and (left column) $\\bm{\\phi}$ and (second column) $\\bm{\\chi}$.", out.width="100%">>= 
@

>>>>

load(file="results/simulations_gdsc_res5.Rdata")
n_eff.mean <- round(mean(summary.mcmc2$summary[, "n_eff"]), 0)
n_eff.min <- round(min(summary.mcmc2$summary[, "n_eff"]), 0)
time.mean <- round((time.mcmc2*1000/n_eff.mean)/3600, 0)
time.min <- round((time.mcmc2*1000/n_eff.min)/(7*24*3600), 0)
@	
  The lasso and ridge models in this section were fit separately per drug, 
  using the \texttt{R} package \texttt{glmnet} 
  \cite[]{friedman_regularization_2010} with cross
	validated penalty parameters.

@@

  For completeness, Figure \ref{fig:simulations_gdsc_res5} displays samples
  from the full Bayesian NIG posterior, as introduced in Section 
  \ref{sec:gibbssampler}. Unfortunately, we were not able to generate enough
  samples for a reliable posterior using the Gibbs procedure as described in 
  the Section. We generated the samples from the full Bayesian posterior using
  stan through the R package \texttt{rstan} 
  \cite[]{guo_rstan:_2018} (see \url{https://github.com/magnusmunch/NIG} for 
  the implemented stan code). The Figure shows that the posteriors for
  $\bm{\phi}$ and $\bm{\chi}$ have unexpected shapes and seem to have explored
  only a small part of the posterior space. This indication that
  posterior inference is not reliable is confirmed by the average and minimum
  effective sample sizes of \Sexpr{n_eff.mean} and \Sexpr{n_eff.min}over all 
  parameters, respectively. For reliable inference,
  \cite{kruschke_doing_2015} recommends an effective sample size of at least 
  10 000. Inspection of the computation times and a quick calculation learns 
  that we have to sample from the posterior
  for at least \Sexpr{time.mean} hours to obtain an average effective sample 
  size of 10 000. Moreover, to achieve a minimum effective sample size of 10 000 
  for all parameters, the sampler has to be run for about \Sexpr{time.min}
  weeks.
<<simulations_gdsc_res5, fig.cap="Simulation results for Scenario 3: Samples from full Bayesian NIG posteriors (see Section \\ref{sec:gibbssampler}) with true value superimposed (red line) for (top row) group 1, (second row) group 2, (third row) group 3, and (bottom row) group 4 for both features and drugs, and (left column) $\\bm{\\phi}$ and (second column) $\\bm{\\chi}$. Note that the red line is not present in most of the Figures, indicating that the posterior is far of from the true value", out.width="100%">>= 
@
